---
date: 2022-10-03
author: "Chao Dai"
format: 
  html:
    code-fold: true
execute: 
  warning: false
  message: false
---


```{r}
#| label: setup
#| include: false

library(tidyverse)
library(data.table)
```


```{r}
#| label: read-in-files
noisy = fread('../code/results/Geuvadis/noise/Geuvadis_perind.counts.noise.gz')
#noisy_intr = fread('../code/results/Geuvadis/noise/Geuvadis_perind.counts.noise_by_intron.gz')
```




The `{out}_perind.counts.noise.gz` noise count file has these main features: 

- noisy coordinates are the min(starts) and max(ends) of all functional introns' coordinates within each cluster
- reads are summed up
  
Effectively, it indicates for each cluster the number of reads attributed to noisy splicing (by sample)
```{r}
split_chrom = function(chrom_v) {
  chrom_v = str_split(chrom_v, ":") %>% 
  map(~ list(chrom = .x[1], 
             start = .x[2], 
             end = .x[3],
             strand = str_extract(.x[4], "[\\+\\-]"),
             clu = str_extract(.x[4], "clu_[0-9]+"),
             noisy = str_detect(.x[4], "\\*")
             )) %>% 
  map(as.data.table) %>% 
  rbindlist()
  return(chrom_v) # data.table
}
```

```{r}
#| label: filter-noisy-only

noisy2 = noisy[, c(split_chrom(chrom), .SD), .SDcols = patterns("^ERR")] %>% 
  melt(measure.vars = patterns("^ERR.+"), 
       variable.name = "samp", 
       value.name = "frac"
       ) %>% 
  .[, `:=`(
    num = str_split(frac, "/", simplify = T)[, 1] %>% as.integer,
    denom = str_split(frac, "/", simplify = T)[, 2] %>% as.integer
  )]
noisy2[, frac := num/denom]
```



Number of clusters having legit noisy introns `29690`
```{r}
noisy2[noisy & num > 2 & denom > 30 & frac > 0 & frac < 1, clu] %>% unique %>% length
```

Number of samples having noisy splicing?

```{r}
#| fig-cap: "ecdf of number of samples detected per noisy splicing event"
noisy2[noisy & num > 2 & denom > 30 & frac > 0 & frac < 1,] %>% 
  .[, .(n_samp = length(unique(samp))), by = clu] %>% 
  ggplot() + stat_ecdf(aes(n_samp)) + 
    geom_hline(yintercept = .8, color = "red", linetype = "dashed")
  
```
Do some samples have more noisy splicing?

```{r}
#| fig-cap: "ecdf of number of noisy splicing events per sample"

noisy2[noisy & num > 2 & denom > 30 & frac > 0 & frac < 1,]  %>% 
  .[, .(n_noisy_clu = length(unique(clu))), by = samp] %>% 
  .[, 
    .(samp = str_split(samp, "\\.", simplify = T)[, 1], 
      n_noisy_clu)
  ] %>% 
  ggplot() + stat_ecdf(aes(n_noisy_clu)) + 
    geom_hline(yintercept = .8, color = 'red', linetype = "dashed")
```

```{r}
#| fig-cap: "Number of splicing events detected per sample"
noisy2[noisy & num > 2 & denom > 30 & frac > 0 & frac < 1,]  %>% 
  .[, .(n_noisy_clu = length(unique(clu))), by = samp] %>% 
  .[, 
    .(samp = str_split(samp, "\\.", simplify = T)[, 1], 
      n_noisy_clu,
      grp = cut_interval(n_noisy_clu, length = 3000, dig.lab = 6))
  ] %>% 
  ggplot() + geom_boxplot(aes(grp, n_noisy_clu)) + 
    geom_point(aes(grp, n_noisy_clu, color = samp), position = position_jitter(width = .2, height = .2)) + 
    theme(legend.position = "None")
```

min: 2522, max 17223 noisy clusters per sample.

```{r}
noisy2[noisy & num > 2 & denom > 30 & frac > 0 & frac < 1,]  %>% 
  .[, .(n_noisy_clu = length(unique(clu))), by = samp] %>% 
  .[, 
    .(samp = str_split(samp, "\\.", simplify = T)[, 1], 
      n_noisy_clu)
  ] %>% 
  .[order(-n_noisy_clu)]
```














