---
params:
  population: "EUR"
title: "Geuvadis sfQTL - EUR"
categories: 
  - QTL
  - Geuvadis
  - "noisy splicing"
  - 'EUR'
date: 2022-12-12
author: "Chao Dai"
---

### Samples used

I used 364 EUR Geuvadis samples that have a match in 1000 genome project. These are originally from Ankeeta's junction files. I noticed all of these samples only have 1 single ERR file, even though there should be many individuals correspond to more than 1 ERR files.

I run EUR and YRI separately.

```{r}
#| label: setup
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(data.table)
library(GenomicFeatures)
library(GenomicRanges)
```

```{r}
metadata = fread("../code/resources/geuvadis-sample-run-pop-lookup.tsv")
```

```{r}
perm_out_cols = c("pid", "pchr", "pstart", "pend", "pstrand", "num_snp_tested",
                  "best_nominal_dist", "best_gid", "best_gchr", "best_gstart",
                  "best_gend", "true_dof", "esti_dof", "beta_ml1", "beta_ml2",
                  "pval_nom", "pval_r2", "regression_slop", "pval_emp", 
                  "pval_adj", "q")
folder = paste0("../code/results/qtl/noisy/Geuvadis/", params$population, "/cis_100000/perm", collapse = "")
qtls = dir(folder, ".+addQval.txt.gz", full.names = T)
qtls = naturalsort::naturalsort(qtls)
names(qtls) = str_extract(qtls, "chr[0-9]{1,2}")
qtls = map(qtls, ~fread(.x, col.names = perm_out_cols))
```

qvalue summaries

::: panel-tabset
#### chr1 - chr5

```{r}
map(qtls[1:5], ~ summary(.x$q))
```

#### chr6 - chr10

```{r}
map(qtls[6:10], ~ summary(.x$q))
```

#### chr11 - chr15

```{r}
map(qtls[11:15], ~ summary(.x$q))
```

#### chr16 - chr22

```{r}
map(qtls[16:22], ~ summary(.x$q))
```
:::

### qqplots - qtltools adjusted p-values

These p-values are adjusted for number of variants, but not adjusted for phenotypes.

::: panel-tabset
#### chr1 - chr5

```{r}
leafcutter::multiqq(
  list(chr1 = qtls$chr1$pval_emp,
       chr2 = qtls$chr2$pval_emp,
       chr3 = qtls$chr3$pval_emp,
       chr4 = qtls$chr4$pval_emp,
       chr5 = qtls$chr5$pval_emp
       )
)
```

#### chr6 - chr10

```{r}
leafcutter::multiqq(
  list(chr6 = qtls$chr6$pval_emp,
       chr7 = qtls$chr7$pval_emp,
       chr8 = qtls$chr8$pval_emp,
       chr9 = qtls$chr9$pval_emp,
       chr10 = qtls$chr10$pval_emp
       )
)
```

#### chr11 - chr15

```{r}
leafcutter::multiqq(
  list(chr11 = qtls$chr11$pval_emp,
       chr12 = qtls$chr12$pval_emp,
       chr13 = qtls$chr13$pval_emp,
       chr14 = qtls$chr14$pval_emp,
       chr15 = qtls$chr15$pval_emp
       )
)
```

#### chr16 - chr20

```{r}
leafcutter::multiqq(
  list(chr16 = qtls$chr16$pval_emp,
       chr17 = qtls$chr17$pval_emp,
       chr18 = qtls$chr18$pval_emp,
       chr19 = qtls$chr19$pval_emp,
       chr20 = qtls$chr20$pval_emp
       )
)
```

#### chr16 - chr20

```{r}
leafcutter::multiqq(
  list(
       chr21 = qtls$chr21$pval_emp,
       chr22 = qtls$chr22$pval_emp
       )
)
```
:::

### How many qtls passing qval \< 0.1?

```{r}
puta_qtls = map(qtls, ~ .x[q < .1]) %>% 
  rbindlist
```

Total number of qtls: `r length(puta_qtls$pid)`

```{r}
chroms = forcats::as_factor(paste("chr", 1:22, sep=""))
puta_qtls[, .(n_qtls = length(unique(pid))), by = pchr] %>% 
  .[, .(qtls = n_qtls, chr = factor(pchr, chroms))] %>% 
  ggplot() + geom_col(aes(chr, qtls)) +
    labs(x = NULL) + 
    theme(axis.text.x = element_text(angle = 30))
```

```{r}
# load in annotation
anno.gene = "../../genome_index/hs38/gencode.v38.primary_assembly.annotation.dataframe.csv"
mycols = c("seqname", "start", "end", "strand", "gene_name")
anno.gene = fread(anno.gene) 
anno.gene = anno.gene[gene_type %in% c("protein_coding") & 
                        feature %in% c("gene") & 
                        source %in% c("HAVANA"), ..mycols] %>% 
  unique %>% 
  makeGRangesFromDataFrame(keep.extra.columns = T)

# remove scaffolds
chroms = paste("chr", 1:22, sep = "")
anno.gene = anno.gene[seqnames(anno.gene) %in% chroms]
```

```{r}

# annotate pid with gene names
puta_loci = puta_qtls[, .(pid, pchr, pstart, pend, pstrand)] %>% 
  makeGRangesFromDataFrame(keep.extra.columns = T, 
                           seqnames.field = "pchr", 
                           start.field = "pstart", 
                           end.field = "pend",
                           strand.field = "pstrand")

hits = GenomicRanges::findOverlaps(puta_loci, anno.gene, minoverlap = 1)
hits = cbind(puta_loci[queryHits(hits)] %>% mcols,
                 anno.gene[subjectHits(hits)] %>% mcols
                 )
hits = as.data.table(hits) %>% 
  .[,  list(genes = pull(.SD, gene_name) %>% paste0(collapse = ",")), by = "pid"]

puta_qtls = hits[puta_qtls, on = "pid", nomatch = NA]

```

```{r}
puta_qtls[!is.na(genes)][order(q)]
```
