---
title: "Geuvadis sfQTL - EUR"
categories:
  - QTL
  - Geuvadis
  - 'EUR'
  - "noisy splicing"
  - 'YRI'
date: 2023-1-5
author: "Chao Dai"
---

# Introduction

  - method 1: uses ranknorm and a few filters to create phenotype bed. Then
subsequent qtl steps are the same.

  - method 2: uses leafcutter prepphenotype python script. Then subsequent QTL 
steps are the same.



```{r}
#| label: setup

suppressMessages(library(data.table))
suppressMessages(library(tidyverse))
```


-----------------------------------------------------------------------

# Geuvadis - EUR - method 1

Compare sfQTL and QTLs called using randomized phenotypes.

```{r}
# QTLs called on randomized noise splicing phenotypes
rqtls = dir("../code/results/qtlQC/perm/Geuvadis/EUR/cis_100000", 
            "chr.*addQ.*gz", full.names = T) %>% naturalsort::naturalsort()
names(rqtls) = str_extract(rqtls, "chr[0-9]{1,2}")
rqtls = map(rqtls, fread)
```

```{r}
# QTLs called on noisy splicing

qtls = dir("../code/results/qtl/noisy/Geuvadis/EUR/cis_100000/perm", 
           "chr.*addQ.*gz", full.names = T) %>% naturalsort::naturalsort()
names(qtls) = str_extract(qtls, "chr[0-9]{1,2}")
qtls = map(qtls, fread)
```

```{r}
#| message: false
#| warning: false
# plot qq for each chromosome
ggls = map(paste("chr", 1:22, sep=""),
            ~ leafcutter::multiqq(list("random" = rqtls[[.x]][["pval_emp"]],
                                       "qtl" = qtls[[.x]][["pval_emp"]])) + 
                lims(x = c(0,4), y = c(0,4)) +
                labs(title = paste0(.x,
                                    " N=", length(qtls[[.x]][["phenotype_id"]])
                                    )) +
                theme(legend.title = element_blank(),
                      legend.position = "bottom",
                      legend.box.spacing = unit(0, "pt"),
                      legend.margin=margin(0,0,0,0),
                      text = element_text(size=8))
            )
```

## qq plots

N = number of phenotypes. Since I only shuffled labels of phenotype ids, the number of phenotype ids in actual data and randomized data are the same.

`pval_emp` or the The P-value of association adjusted for the number of variants tested in cis given by the direct method is plotted. Since it's empirical p-value, with 1000 permutations, p value saturates at 0.001.

::: panel-tabset
#### chr1-6

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[1:6], nrow = 2)
```

#### chr7-12

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[7:12], nrow = 2)
```

#### chr13-18

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[13:18], nrow = 2)
```

#### chr19-22

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[19:22], ncol = 3)
```
:::

## Number of qtls at 10% FDR

Number of phenotypes: 

  - `r map_int(qtls, ~.x[["phenotype_id"]] %>% length) %>% sum`

Number of qtls at 10% FDR: 

  - total: `r map_int(qtls, ~.x[q<0.1] %>% .[["phenotype_id"]] %>% length) %>% sum` - by chrom

```{r}
map_int(qtls, ~.x[q<0.1] %>% .[["phenotype_id"]] %>% length) %>% 
  sort(decreasing = T) %>% print
```



-----------------------------------------------------------------------

# Geuvadis - EUR - method 2

Compare sfQTL and QTLs called using randomized phenotypes.


```{r}
# QTLs called on noisy splicing

qtls2 = dir("../code/results/qtl/noisy/Geuvadis/EUR/cis_100000/perm/method2", 
           "chr.*addQ.*gz", full.names = T) %>% naturalsort::naturalsort()
names(qtls2) = str_extract(qtls2, "chr[0-9]{1,2}")
qtls2 = map(qtls2, fread)
```

```{r}
#| message: false
#| warning: false
# plot qq for each chromosome
ggls = map(paste("chr", 1:22, sep=""),
            ~ leafcutter::multiqq(list("random" = rqtls[[.x]][["pval_emp"]],
                                       "qtl" = qtls2[[.x]][["pval_emp"]])) + 
                lims(x = c(0,4), y = c(0,4)) +
                labs(title = paste0(.x,
                                    " N=", length(qtls2[[.x]][["phenotype_id"]])
                                    )) +
                theme(legend.title = element_blank(),
                      legend.position = "bottom",
                      legend.box.spacing = unit(0, "pt"),
                      legend.margin=margin(0,0,0,0),
                      text = element_text(size=8))
            )
```

## qq plots

N = number of phenotypes. Since I only shuffled labels of phenotype ids, the number of phenotype ids in actual data and randomized data are the same.

`pval_emp` or the The P-value of association adjusted for the number of variants tested in cis given by the direct method is plotted. Since it's empirical p-value, with 1000 permutations, p value saturates at 0.001.

::: panel-tabset
#### chr1-6

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[1:6], nrow = 2)
```

#### chr7-12

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[7:12], nrow = 2)
```

#### chr13-18

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[13:18], nrow = 2)
```

#### chr19-22

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[19:22], ncol = 3)
```
:::

## Number of qtls at 10% FDR

Number of phenotypes: 

  - `r map_int(qtls2, ~.x[["phenotype_id"]] %>% length) %>% sum`

Number of qtls at 10% FDR: 

  - total: `r map_int(qtls2, ~.x[q<0.1] %>% .[["phenotype_id"]] %>% length) %>% sum` - by chrom

```{r}
map_int(qtls2, ~.x[q<0.1] %>% .[["phenotype_id"]] %>% length) %>% 
  sort(decreasing = T) %>% print
```


## Method 1 QTLs vs. Method 2 QTLs


```{r}
n1 = map_dfr(qtls, ~.x[q<0.1]) %>% .[, phenotype_id] %>% unique
n2 = map_dfr(qtls2, ~.x[q<0.1]) %>% .[, phenotype_id] %>% unique

```

Number of QTLs found in both methods as a percentage of method 1 QTLs: 

  - `r length(intersect(n1,n2)) / length(unique(n1))`

------------------------------------------------------------------------

# Geuvadis - YRI - method 1


Compare sfQTL and QTLs called using randomized phenotypes.

```{r}
# QTLs called on randomized noise splicing phenotypes
rqtls = dir("../code/results/qtlQC/perm/Geuvadis/YRI/cis_100000", 
            "chr.*addQ.*gz", full.names = T) %>% naturalsort::naturalsort()
names(rqtls) = str_extract(rqtls, "chr[0-9]{1,2}")
rqtls = map(rqtls, fread)
```

```{r}
# QTLs called on noisy splicing

qtls = dir("../code/results/qtl/noisy/Geuvadis/YRI/cis_100000/perm", 
           "chr.*addQ.*gz", full.names = T) %>% naturalsort::naturalsort()
names(qtls) = str_extract(qtls, "chr[0-9]{1,2}")
qtls = map(qtls, fread)
```

```{r}
#| message: false
#| warning: false
# plot qq for each chromosome
ggls = map(paste("chr", 1:22, sep=""),
            ~ leafcutter::multiqq(list("random" = rqtls[[.x]][["pval_emp"]],
                                       "qtl" = qtls[[.x]][["pval_emp"]])) + 
                lims(x = c(0,4), y = c(0,4)) +
                labs(title = paste0(.x,
                                    " N=", length(qtls[[.x]][["phenotype_id"]])
                                    )) +
                theme(legend.title = element_blank(),
                      legend.position = "bottom",
                      legend.box.spacing = unit(0, "pt"),
                      legend.margin=margin(0,0,0,0),
                      text = element_text(size=8))
            )
```

## qq plots

N = number of phenotypes. Since I only shuffled labels of phenotype ids, the number of phenotype ids in actual data and randomized data are the same.

`pval_emp` or the The P-value of association adjusted for the number of variants tested in cis given by the direct method is plotted. Since it's empirical p-value, with 1000 permutations, p value saturates at 0.001.

::: panel-tabset
#### chr1-6

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[1:6], nrow = 2)
```

#### chr7-12

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[7:12], nrow = 2)
```

#### chr13-18

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[13:18], nrow = 2)
```

#### chr19-22

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[19:22], ncol = 3)
```
:::

## Number of qtls at 10% FDR

Number of phenotypes:

-   `r map_int(qtls, ~.x[["phenotype_id"]] %>% length) %>% sum`

Number of qtls at 10% FDR:

-   total: `r map_int(qtls, ~.x[q<0.1] %>% .[["phenotype_id"]] %>% length) %>% sum` - by chrom

```{r}
map_int(qtls, ~.x[q<0.1] %>% .[["phenotype_id"]] %>% length) %>% 
  sort(decreasing = T) %>% print
```







------------------------------------------------------------------------

# Geuvadis - YRI - method 2


Compare sfQTL and QTLs called using randomized phenotypes.


```{r}
# QTLs called on noisy splicing

qtls2 = dir("../code/results/qtl/noisy/Geuvadis/YRI/cis_100000/perm/method2", 
           "chr.*addQ.*gz", full.names = T) %>% naturalsort::naturalsort()
names(qtls2) = str_extract(qtls2, "chr[0-9]{1,2}")
qtls2 = map(qtls2, fread)
```

```{r}
#| message: false
#| warning: false
# plot qq for each chromosome
ggls = map(paste("chr", 1:22, sep=""),
            ~ leafcutter::multiqq(list("random" = rqtls[[.x]][["pval_emp"]],
                                       "qtl" = qtls2[[.x]][["pval_emp"]])) + 
                lims(x = c(0,4), y = c(0,4)) +
                labs(title = paste0(.x,
                                    " N=", length(qtls2[[.x]][["phenotype_id"]])
                                    )) +
                theme(legend.title = element_blank(),
                      legend.position = "bottom",
                      legend.box.spacing = unit(0, "pt"),
                      legend.margin=margin(0,0,0,0),
                      text = element_text(size=8))
            )
```

## qq plots

N = number of phenotypes. Since I only shuffled labels of phenotype ids, the number of phenotype ids in actual data and randomized data are the same.

`pval_emp` or the The P-value of association adjusted for the number of variants tested in cis given by the direct method is plotted. Since it's empirical p-value, with 1000 permutations, p value saturates at 0.001.

::: panel-tabset
#### chr1-6

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[1:6], nrow = 2)
```

#### chr7-12

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[7:12], nrow = 2)
```

#### chr13-18

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[13:18], nrow = 2)
```

#### chr19-22

```{r}
#| warning: false
gridExtra::grid.arrange(grobs = ggls[19:22], ncol = 3)
```
:::

## Number of qtls at 10% FDR

Number of phenotypes:

-   `r map_int(qtls2, ~.x[["phenotype_id"]] %>% length) %>% sum`

Number of qtls at 10% FDR:

-   total: `r map_int(qtls2, ~.x[q<0.1] %>% .[["phenotype_id"]] %>% length) %>% sum` - by chrom

```{r}
map_int(qtls2, ~.x[q<0.1] %>% .[["phenotype_id"]] %>% length) %>% 
  sort(decreasing = T) %>% print
```


## Method 1 QTLs vs. Method 2 QTLs


```{r}
n1 = map_dfr(qtls, ~.x[q<0.1]) %>% .[, phenotype_id] %>% unique
n2 = map_dfr(qtls2, ~.x[q<0.1]) %>% .[, phenotype_id] %>% unique

```

Number of QTLs found in both methods as a percentage of method 1 QTLs: 

  - `r length(intersect(n1,n2)) / length(unique(n1))`





