---
title: "Summarize Geuvadis samples"
description: >
  "Investigate which geuvadis samples have an overlapping genotype file in 1KGP Phase 3 vcf. And which of such samples have 1 or 1+ run ids (ERR)."
categories: 
  - "geuvadis"
  - "1000 genome project"
  - "metadata"
date: 2022-10-20
author: 'Chao Dai'
---

```{r}
#| label: setup
#| include: false
#| message: false
#| warning: false

library(tidyverse)
library(data.table)
```

### Geuvadis Samples

```{r}
# get geuvadis metadata
geuvadis_meta = "../code/resources/geuvadis-sample-run-pop-lookup.tsv"
geuvadis_meta = fread(geuvadis_meta)
```

::: panel-tabset
#### by population

```{r}
#| echo: false
geuvadis_meta[, .(N = length(unique(Sample))), by = "Pop_id"]
```

#### by number of ERRs

```{r}
#| echo: false
geuvadis_meta[, .(Sample, run_id)] %>% 
  unique %>% 
  .[, .(N_ERR = length(unique(run_id))), by = Sample] %>% 
  .[order(-N_ERR)] %>% 
  ggplot() + geom_histogram(aes(N_ERR), binwidth = 1, color = 'white') +
    scale_x_continuous(name = "Number of ERR IDs per sample", breaks = 1:8, labels = 1:8) +
    labs(title = "Geuvadis - all samples")
```
:::

### Common samples between Geuvadis and 1000 genome project genotypes

Sample IDs, aka. `HG#####` or `NA#####` numbers are common keys that link samples
in Geuvadis and 1000 genome project.

Below table shows number of sampels in Geuvadis and 1KGP as well as the number of
samples that can be linked.


```{r}
# get all genotype sample names
cmd = "  view -h /project2/yangili1/zpmu/1kg_b38/CCDG_14151_B01_GRM_WGS_2020-08-05_chr1.filtered.shapeit2-duohmm-phased.vcf.gz | awk '$1 ~ /#CHROM/'"

SamIDs_1kgp = system2("bcftools", cmd, stdout = T) %>% 
  scan(text = ., what = character())
SamIDs_1kgp = SamIDs_1kgp[10:length(SamIDs_1kgp)]

# samples that can be linked between Geuvadis and 1KGP
SamIDs_common = intersect(SamIDs_1kgp, geuvadis_meta$Sample)
```

|                      | N                                        |
|----------------------|------------------------------------------|
| Geuvadis Samples     | `r length(unique(geuvadis_meta$Sample))` |
| 1KGP Phase 3 Samples | `r length(unique(SamIDs_1kgp))`          |
| Samples in both      | `r length(SamIDs_common)`                |


### Focusing on common samples

```{r}
# geuvadis linkable samples that only have 1 ERR IDs
geuvadis_singleERR_samples = geuvadis_meta[Sample %in% SamIDs_common, .(Sample, run_id)] %>% 
  unique %>% 
  .[, .(N_ERR = length(unique(run_id))), by = Sample] %>%
  .[N_ERR == 1, Sample]
```

Out of all `r length(unique(geuvadis_meta$Sample))` Geuvadis samples, 
`r length(geuvadis_singleERR_samples)` of them has 1-to-1 ERR IDs. The other
samples have multiple ERRs per sample.

For my test analysis, I am running only on these 1-to-1 samples to establish 
analysis pipeline, but later on formal analysis will include all linkable samples.


### Saving sample ids to files

Save all linkable Geuvadis sample ids to a text file, with 1 ID per row.
```{r}
#write_lines(SamIDs_common, "../code/resources/geuvadis-1kgp-common-sample-id.txt")
```

Save all 1-to-1 linkable Geuvadis sample ids to a text file, 1 ID per row.
```{r}
#write_lines(geuvadis_singleERR_samples, "../code/resources/geuvadis-1kgp-common-1to1only-sample-id.txt")
```


### Check junction file samples from Ankeeta

```{r}
noise = fread("../code/results/Geuvadis/noise/Geuvadis_perind.counts.noise.gz",
              colClasses = 'character', nrows = 5)
noise_ERRs = names(noise)
noise_ERRs = noise_ERRs[noise_ERRs != "chrom"]
noise_ERRs = str_extract(noise_ERRs, "ERR[0-9]{5,7}")
noise_samples = filter(geuvadis_meta, run_id %in% noise_ERRs) %>% 
  pull(Sample) %>%
  unique
```


These samples in Ankeeta's dataset do not have genotype in 1KGP
```{r}
setdiff(noise_samples, SamIDs_1kgp)
```


Ankeeta's dataset has 452 ERR IDs, smaller than the full Geuvadis dataset (`r length(unique(geuvadis_meta$run_id))`).
37% of samples have more than 1 runs.

```{r}
filter(geuvadis_meta, Sample %in% noise_samples) %>% 
  select(Sample, run_id, Pop_id) %>% 
  group_by(Sample) %>% 
  mutate(N = length(unique(run_id))) %>% 
  mutate(morethan1runs = N > 1) %>% 
  ungroup %>% group_by(morethan1runs) %>% 
  summarise(N_samples = length(unique(Sample)))
```





