---
title: "Geuvadis noisy splicing QTL"
categories: 
  - QTL
  - Geuvadis
  - "noisy splicing"
date: 2022-10-23
author: "Chao Dai"
---

### Samples used

274 Geuvaids samples are used (see [details](Geuvadis-Samples.pub.qmd)) in this 
test run. These are samples that have single run (single ERR ID) that have matching 
genotype in 1000 genome project.
<br>
Ankeeta's dataset has total 452 runs, representing lesst han 452 samples.



```{r}
#| label: setup
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(data.table)
library(GenomicFeatures)
library(GenomicRanges)
```

```{r}
metadata = fread("../code/resources/geuvadis-sample-run-pop-lookup.tsv")
```


```{r}
perm_out_cols = c("pid", "pchr", "pstart", "pend", "pstrand", "num_snp_tested",
                  "best_nominal_dist", "best_gid", "best_gchr", "best_gstart",
                  "best_gend", "true_dof", "esti_dof", "beta_ml1", "beta_ml2",
                  "pval_nom", "pval_r2", "regression_slop", "pval_emp", 
                  "pval_adj", "qval")
qtls = dir("../code/results/qtltools/Geuvadis/cis_100000/perm", 
           ".+addQval.txt.gz", full.names = T)
qtls = naturalsort::naturalsort(qtls)
names(qtls) = str_extract(qtls, "chr[0-9]{1,2}")
qtls = map(qtls, ~fread(.x, col.names = perm_out_cols))
```

qvalue summaries

::: {.panel-tabset}
#### chr1 - chr5
```{r}
map(qtls[1:5], ~ summary(.x$qval))
```
#### chr6 - chr10
```{r}
map(qtls[6:10], ~ summary(.x$qval))
```

#### chr11 - chr15
```{r}
map(qtls[11:15], ~ summary(.x$qval))
```
#### chr16 - chr22
```{r}
map(qtls[16:22], ~ summary(.x$qval))
```

:::



### qqplots - qtltools adjusted p-values

These p-values are adjusted for number of variants, but not adjusted for phenotypes.

:::{.panel-tabset}
#### chr1 - chr5
```{r}
leafcutter::multiqq(
  list(chr1 = qtls$chr1$pval_emp,
       chr2 = qtls$chr2$pval_emp,
       chr3 = qtls$chr3$pval_emp,
       chr4 = qtls$chr4$pval_emp,
       chr5 = qtls$chr5$pval_emp
       )
)
```
#### chr6 - chr10
```{r}
leafcutter::multiqq(
  list(chr6 = qtls$chr6$pval_emp,
       chr7 = qtls$chr7$pval_emp,
       chr8 = qtls$chr8$pval_emp,
       chr9 = qtls$chr9$pval_emp,
       chr10 = qtls$chr10$pval_emp
       )
)
```
#### chr11 - chr15
```{r}
leafcutter::multiqq(
  list(chr11 = qtls$chr11$pval_emp,
       chr12 = qtls$chr12$pval_emp,
       chr13 = qtls$chr13$pval_emp,
       chr14 = qtls$chr14$pval_emp,
       chr15 = qtls$chr15$pval_emp
       )
)
```

#### chr16 - chr20
```{r}
leafcutter::multiqq(
  list(chr16 = qtls$chr16$pval_emp,
       chr17 = qtls$chr17$pval_emp,
       chr18 = qtls$chr18$pval_emp,
       chr19 = qtls$chr19$pval_emp,
       chr20 = qtls$chr20$pval_emp
       )
)
```


#### chr16 - chr20
```{r}
leafcutter::multiqq(
  list(
       chr21 = qtls$chr21$pval_emp,
       chr22 = qtls$chr22$pval_emp
       )
)
```

:::


### How many qtls passing qval < 0.1?

```{r}
puta_qtls = map(qtls, ~ .x[qval < .1]) %>% 
  rbindlist
```

Total number of qtls: `r length(puta_qtls$pid)`

```{r}
chroms = forcats::as_factor(paste("chr", 1:22, sep=""))
puta_qtls[, .(n_qtls = length(unique(pid))), by = pchr] %>% 
  .[, .(qtls = n_qtls, chr = factor(pchr, chroms))] %>% 
  ggplot() + geom_col(aes(chr, qtls)) +
    labs(x = NULL) + 
    theme(axis.text.x = element_text(angle = 30))
```



```{r}
# load in GFF
gff = "../../genome_index/hs38/gencode.v38.annotation.gff3"

txdb = maketx
```


