---
title: "Validate GTEx V8 junction file"
categories: 
  - GTEX
  - "junctions"
date: 2022-10-20
author: Chao Dai
---

GTEx V8 junction file was downloaded from [GTEx portal](https://gtexportal.org/home/datasets), specifically via this link: [GTEx_Analysis_2017-06-05_v8_STARv2.5.3a_junctions.gct.gz](https://storage.googleapis.com/gtex_analysis_v8/rna_seq_data/GTEx_Analysis_2017-06-05_v8_STARv2.5.3a_junctions.gct.gz).

Downloaded file path: 

- `resources/GTEx/juncs/GTEx_Analysis_2017-06-05_v8_STARv2.5.3a_junctions.gct.gz`

Because the file is rather large, a python script is used to compute the histogram of each junction. 

- `workflow/scripts/make_histogram_of_GTEx_juncs.py`


```{r}
#| label: setup
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(data.table)

```



### The GTEx V8 junction file include low reads

See plots below. It seems this data set does not filter out 
```{r}
# histograms for each junction
hist_data = fread("../code/resources/GTEx/juncs/GTEx_Analysis_2017-06-05_v8_STARv2.5.3a_junctions_histogram.txt")
# junction coordinates and gene id
col_data = fread("../code/resources/GTEx/juncs/GTEx_Analysis_2017-06-05_v8_STARv2.5.3a_junctions_names.txt", col.names = c("junc", "gene_id"))
```

:::{.panel-tabset}

#### Samples with 0-5 reads
```{r}
ggplot(hist_data) + geom_histogram(aes(`bin_0-5`), bins=20) +
  labs(x = "Num. of samples with <5 reads", y = NULL)
```

#### Distribution of Min(reads)
```{r}
ggplot(hist_data) + stat_ecdf(aes(Min), bins = 20) +
  labs(x = "minimum reads", y = NULL)
```
:::

